\documentclass[11pt,a4paper]{article}

% ---- Packages ----
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{float}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, shapes.geometric, fit, calc}
\usepackage{listings}

% ---- Listings style ----
\lstset{
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  frame=single,
  backgroundcolor=\color{gray!10},
  keywordstyle=\color{blue},
  commentstyle=\color{green!50!black},
  stringstyle=\color{red!70!black},
}

% ---- Custom commands ----
\newcommand{\best}[1]{\textbf{#1}}

% ---- Title ----
\title{
  \vspace{-1cm}
  \LARGE\textbf{Multimodal Presurgical Brain Tumor Classification\\
  Using Statistical Machine Learning}
  \vspace{0.3cm}
}
\author{
  \textbf{Group 12} \\
  STAT3612: Statistical Machine Learning (Spring 2026)
}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
This report presents our approach to presurgical brain tumor classification using multimodal data. We design an ensemble system that combines radiology report text (via BioClinicalBERT and TF-IDF), clinical features (signal intensity patterns, tumor location), and MRI-derived features (ResNet-50 embeddings with per-modality PCA) to classify brain tumors into five types: Glioma, Meningioma, Brain Metastase, Sellar region tumors, and Pineal/Choroid Plexus tumors.

Our design is guided by a systematic analysis of each data modality's discriminative power. We find that radiology reports alone achieve $\sim$84\% Micro-F1 and provide the dominant signal, while clinical features ($\sim$67\%), image features ($\sim$57\%), and raw radiomic features ($\sim$52\%, near random) contribute progressively less. Based on this analysis, we design a dual-path neural network where BioClinicalBERT processes report text and a tabular encoder handles clinical and location features, connected by a cross-modal gate that adaptively controls information flow between paths. We complement this with three gradient-boosted tree models (XGBoost, LightGBM, CatBoost) trained on TF-IDF report features combined with clinical data. The final 6-model weighted ensemble achieves \best{87.25\%} out-of-fold Micro-F1 on 2,266 training samples across 5-fold cross-validation.

A central challenge is the potential for distribution shift between training and test sets. We employ NN-dominant ensemble weighting and hierarchical location encoding as potential mitigation strategies, and note that OOF estimates may be optimistic if test distributions differ from training.
\end{abstract}

\tableofcontents
\newpage

% ============================================================
% SECTION 1: Introduction
% ============================================================
\input{sections/01_introduction}

% ============================================================
% SECTION 2: Dataset & EDA
% ============================================================
\input{sections/02_dataset}

% ============================================================
% SECTION 3: Methodology
% ============================================================
\input{sections/03_methodology}

% ============================================================
% SECTION 4: Experiments & Results
% ============================================================
\input{sections/04_experiments}

% ============================================================
% SECTION 5: Discussion
% ============================================================
\input{sections/05_discussion}

% ============================================================
% SECTION 6: Conclusion
% ============================================================
\input{sections/06_conclusion}

% ============================================================
% References
% ============================================================
\bibliographystyle{unsrt}
\begin{thebibliography}{10}

\bibitem{price2024}
Price M, Ballard C, Benedetti J, et al.
CBTRUS statistical report: primary brain and other central nervous system tumors diagnosed in the United States in 2017--2021[J].
\textit{Neuro-oncology}, 2024, 26(Suppl 6): vi1.

\bibitem{wang2024}
Wang Y R J, Wang P, Yan Z, et al.
Advancing presurgical non-invasive molecular subgroup prediction in medulloblastoma using artificial intelligence and MRI signatures[J].
\textit{Cancer Cell}, 2024, 42(7): 1239--1257.

\bibitem{he2016}
He K, Zhang X, Ren S, et al.
Deep residual learning for image recognition[C].
\textit{Proceedings of the IEEE CVPR}, 2016: 770--778.

\bibitem{devlin2019}
Devlin J, Chang M, Lee K, Toutanova K.
BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding.
\textit{arXiv preprint arXiv:1810.04805}, 2019.

\bibitem{alsentzer2019}
Alsentzer E, Murphy J, Boag W, et al.
Publicly Available Clinical BERT Embeddings.
\textit{arXiv preprint arXiv:1904.03323}, 2019.

\bibitem{chen2016}
Chen T, Guestrin C.
XGBoost: A Scalable Tree Boosting System.
\textit{Proceedings of the 22nd ACM SIGKDD}, 2016: 785--794.

\bibitem{ke2017}
Ke G, Meng Q, Finley T, et al.
LightGBM: A Highly Efficient Gradient Boosting Decision Tree.
\textit{Advances in NeurIPS}, 2017: 3146--3154.

\bibitem{prokhorenkova2018}
Prokhorenkova L, Gusev G, Vorobev A, et al.
CatBoost: Unbiased Boosting with Categorical Features.
\textit{Advances in NeurIPS}, 2018.

\bibitem{lin2017}
Lin T Y, Goyal P, Girshick R, et al.
Focal Loss for Dense Object Detection.
\textit{Proceedings of the IEEE ICCV}, 2017: 2980--2988.

\end{thebibliography}

\end{document}
